\chapter{The hotplug subsystem}
\section{udev}

The {\tt /dev} directory under Frugalware is a ramdisk and every device node is created automatically during the system boot by the hotplug subsystem and especially by \textit{udev}. This means there will not be any unnecessary device nodes in /dev, and but means if you create a device node manually, it will extists only till next shutdown only.

If you want to create cusrom rules, place a \textit{rule file} under
{\tt /etc/udev/rules.d/}. It is also recommended to start the filename with a
number less than 40 (for example {\tt 35-my.rules} is correct) so that your
settings will take predence. Alternaticely, you can force Frugalware to create
a missing device node "manually" during each boot by editing the
{\tt /etc/sysconfig/udev} configuration file.

{\tt udev} requires sysfs, so that it will only work with the kernel series
2.6.x only. Do \textit{not} try to run Frugalware with kernel series 2.4.x with
udev.

\section{Pen/Thumb drivers}

Pendrives (also known as thumbdrives) are well-supported through the hotplug scripts and udev. If you insert a pendrive to your USB slot, udev will create it's device node in {\tt /dev}. Most pendrives contain only one partition and it's filesystem is {\tt vfat}. In most cases the pendrive will behave like a scsi disc. This means you can access the pendrive as {\tt /dev/sda} and it's first partition as {\tt /dev/sda1}. Placing the following line:

{\tt /dev/sda1        /mnt/pendrive    auto        defaults,noauto,user 0 0}

will let users mount their pendrive if the device node exists (if they are inserted it to the slot).

\section{Digital cameras}

Typically there are two types of digital cameras currently exists. First, lots of cameras behave like a pendrive, you can mount them and easily copy the pictures from them.

Other type of cameras support the \textit{Picture Transfer Protocol (PTP)}.
You can grab the pictures from them (and do lots of other actions) with
{\tt gphoto2}. (If it's not available on your system, a single
{\tt sudo pacman -S gphoto2} will install it onto your system.) If you like
graphical frontends, you are looking for {\tt gtkam} (uses GTK+) or
{\tt digikam} (uses QT).

\section{Automounting: {\tt D-BUS}, {\tt HAL} and {Ivman} (newschool)}

{\tt D-BUS} is a simple IPC library based on messages. {\tt HAL} is a hardware
abstraction layer which uses {\tt D-BUS}. {\tt Ivman} is based on {\tt HAL} and
uses {\tt pmount} ("policy mount"), which is a wrapper around the standard
mount program which permits normal users to mount removable devices without a
matching {\tt /etc/fstab} entry.

{\tt Ivman} is a daemon to automount cdrom's and dvd's when inserted in a
drive, or autostart playing audio cd's or video dvd's. It is 100\% userspace,
so it is a safe replacement for {\tt submount}.

If you want to change the default settings, all config files are located in
{\tt /etc/ivman}. They are plain xml files, just read them, everything is
quite self-explaining.

\section{Automounting: {\tt submount} (oldschool)}

This method is a bit obsolete, you will have to install submount from the
{\tt extra} repo with {\tt sudo pacman -S submount}.

The {\tt submount} introduces a new filesystem, called {\tt subfs}. Its aim is to prevent users having to manually mount and umount their removable media (CD, floppy, etc). In Frugalware you can configure what level of automatism you want submount to do for you. 

By default (this is the {\tt auto} mode) submount detects your CD/DVD and floppy drives, makes symlink for them in {\tt /dev}, removes old entries and puts new ones to {\tt /etc/fstab} and mounts drives with subfs, so the users don't have to mount or umount drives manually.

If you don't like this automatism, you can choose from the following modes:

The {\tt dirs} mode: deletes floppy,cdrom,dvd,cdrecorder,dvdrecorder dirs in {\tt /media} and creates which is necessary, also creates entries in {\tt /etc/fstab}, but you have to mount devices manually.

The {\tt symlinks} mode: creates floppy,cdrom,dvd,cdrecorder,dvdrecorder symlinks if necessary in {\tt /dev}, but nothing more.

The {\tt manual} mode: do nothing, like in good old Slackware.

You can set submount's mode in the {\tt /etc/sysconfig/submount} configuration file. Also there you can configure extra mount options like {\tt iocharset} or {\tt codepage}, etc.
