\chapter{X Window System}

\section{Configuring your graphics card}
\label{section:xconfig}

If you install X, setup will automatically run {\tt xconfig}, our X configuration utility. First, xconfig will detect your configuration and will create a simple configuration file. Then it will ask you about the screen resolution and depth. Then, it will create the real configuration file with the following extras:

\begin{itemize}
\item Automatically configures mouse wheel, if found
\item Sets keyboard layout, according to the selected language
\item Enables DPMS
\item Allows users to use the DRI extension
\end{itemize}

After generating the config file it will start the X server and you will have to click to the {\tt OK} button, so {\tt xconfig} will know that the configuration was really successfull.

\section{3D acceleration (and binary drivers)}

If X has built-in 3d acceleration support for your card, then {\tt xconfig} will write the necessary entries to {\tt /etc/X11/xorg.conf} and X will load those module(s). 

If you have an ATI or NVidia card, then you will probably need binary drivers. Obtaining NVidia binary driver is fairly simple:

{\tt vmiklos@vmhome:~\$ su -c 'pacman -Sy nvidia'}

\textbf{Warning:} you'll have to re-run {\tt pacman -S nvidia} after each kernel update!

If you have an ATI card, then you'll have to install it manually. Currently we don't have any hardware to create and test an ATI binary driver package. If you can (temporarily) donate us such a card, feel free to write us!
